---
title: "T7aed"
author: "Rafael Cabrera, Juan Camilo Rodriguez, Guillermo Ribero, David Melendez"
date: '2022-05-13'
output: html_document
---

#### 1. Consider the two data sets

$$
\mathbf{X}_{1}=\left[\begin{array}{ll}
3 & 7 \\
2 & 4 \\
4 & 7
\end{array}\right] \text { and } \mathbf{X}_{2}=\left[\begin{array}{ll}
6 & 9 \\
5 & 7 \\
4 & 8
\end{array}\right]
$$

##### for which

$$
\overline{\mathbf{x}}_{1}=\left[\begin{array}{l}
3 \\
6
\end{array}\right], \quad \overline{\mathbf{x}}_{2}=\left[\begin{array}{l}
5 \\
8
\end{array}\right]
$$

##### and

$$
\mathbf{S}_{\text {pooled }}=\left[\begin{array}{ll}
1 & 1 \\
1 & 2
\end{array}\right]
$$

#### (a) Calculate the linear discriminant function

$\hat{y}=\left(\overline{\mathbf{x}}{1}-\overline{\mathbf{x}}{2}\right)^{\prime} \mathbf{S}_{\text {pooled }}^{-1} \mathbf{x}=\hat{\mathbf{a}}^{\prime} \mathbf{x}$.


```{r}
X1_mean = c(3,6)
  
X2_mean = c(5,8)

Spooled = cbind(c(1,1),c(1,2))

y_gorrito = t(X1_mean-X2_mean)%*%solve(Spooled)
```
$$
\hat{y} = (-2,0) X_0
$$


##### (b) Classify the observation

$\mathbf{x}{0}^{\prime}=\left[\begin{array}{cc}2 & 7\end{array}\right]$ as population $\pi{1}$ or population $\pi_{2}$, 

##### using The Estimated Minimum ECM Rule for Two Normal Populations with equal priors and equal costs.

```{r}
X_0 = c(2,7)
Minimizan_ECM = y_gorrito %*% X_0 - (1/2)*y_gorrito%*%(X1_mean+X2_mean)

```
##### Como $4>0$ clásificamos la observación $X_0$ en $\pi_1$

##### (c) Develop a linear classification function for the data using $\hat{y}$ from (a).

```{r}
y_gorrito_0 = y_gorrito%*%X_0

```

##### (d) Using the function in (c) and

$\hat{m}=\frac{1}{2}\left(\overline{\mathbf{x}}{1}-\overline{\mathbf{x}}{2}\right)^{\prime} \mathbf{S}{\text {pooled }}^{-1}\left(\overline{\mathbf{x}}{1}+\overline{\mathbf{x}}_{2}\right)$, construct the "confusion matrix" by classifying the given observations.

```{r}
m_gorro = (1/2) * y_gorrito%*%(X1_mean+X2_mean)

#Clasificando Dataset X1
X1 = cbind(c(3,2,4),c(7,4,7))
n1c = 0 #clasificaciones correctas
n1m = 0 #clasificaciones incorrectas
for  (i in c(1,2,3)){
  print(i)
  y_gorrito_0 = y_gorrito%*%X1[i,]
  regla_fisher = y_gorrito_0 - m_gorro
  if (regla_fisher>=0){ #clasifica en pie1
   n1c=n1c+1
  }else{#clasifica en pie2
   n1m=n1m+1 
  }
}
print(n1c)
print(n1m)

#Clasificando Dataset X2
X2 = cbind(c(6,5,4),c(9,7,8))
n2c = 0 #clasificaciones correctas
n2m = 0 #clasificaciones incorrectas
for  (i in c(1,2,3)){
  print(i)
  y_gorrito_0 = y_gorrito%*%X2[i,]
  regla_fisher = y_gorrito_0 - m_gorro
  if (regla_fisher>=0){ #clasifica en pie2
   n2c=n2c+1
  }else{#clasifica en pie1
   n2m=n2m+1 
  }
}
print(n2c)
print(n2m)

#Matriz de confusion
confusion_matrix = cbind(c(n1c,n2m),c(n1m,n2c))

```

Entonces clasificamos $X_0$ según la regla de Fisher como $\pi_1$ y la matriz de confusíon es: 
 


| <span>            |             |               |
|-------------------|-------------|---------------|
| $\pi_1$           | 3           | 0             |
| $\pi_2$           | 2           | 1             |

##### (e) Given the results in (d), calculate the apparent error rate (APER).

```{r}
APER = (n1m+n2m)/(6);APER
```

#### 2. Suppose that $n_{1}=11$ and $n_{2}=12$ observations are made on two random variables $X_{1}$ and $X_{2}$, where $X_{1}$ and $X_{2}$ are assumed to have a bivariate normal distribution with a common covariance matrix $\boldsymbol{\Sigma}$, but possibly different mean vectors $\boldsymbol{\mu}{1}$ and $\boldsymbol{\mu}{2}$ for the two samples. The sample mean vectors and pooled covariance matrix are

$$
\begin{array}{l}
\overline{\mathbf{x}}_{1}=\left[\begin{array}{l}
-1 \\
-1
\end{array}\right] ; \quad \overline{\mathbf{x}}_{2}=\left[\begin{array}{l}
2 \\
1
\end{array}\right] \\
\mathbf{S}_{\text {pooled }}=\left[\begin{array}{rr}
7.3 & -1.1 \\
-1.1 & 4.8
\end{array}\right]
\end{array}
$$

##### (a) Construct Fisher's (sample) linear discriminant function:

$\hat{y}=\left(\overline{\mathbf{x}}{1}-\overline{\mathbf{x}}{2}\right)^{\prime} \mathbf{S}_{\text {pooled }}^{-1} \mathbf{x}=$ $\hat{\mathbf{a}}^{\prime} \mathbf{x}$.

```{r}
n1 = 11
n2 = 12
x1_barra = c(-1,-1)
x2_barra = c(2,1)
Spooled = cbind(c(7.3,-1.1),c(-1.1,4.8))
y_gorrito = t(x1_barra-x2_barra)%*%solve(Spooled)

```

##### (b) Assign the observation $\mathbf{x}{0}^{\prime}=\left[\begin{array}{ll}0 & 1\end{array}\right]$ to either population $\pi{1}$ or $\pi_{2}$. Assume equal costs and equal prior probabilities.

```{r}
X_0 = c(0,1)
m_gorro = (1/2) * y_gorrito%*%(x1_barra+x2_barra)
y_gorrito_0 = y_gorrito%*%X_0
regla_fisher = y_gorrito_0 - m_gorro
```

Concluimos que la observación $X_0$ corresponde a $\pi_2$ por la regla de fisher